A quantitative framework for investigating the reliability of empirical network construction

### 2019-02-21 manuscript/reply_to_reviewers_2.tex

		\begin{refquote}

			I think the authors miss one key discomfort with empirical bayes. It seems vaguely anti-ecology, because it assumes that we can L289 : “use information from abundant and easily-sampled species to predict interaction probabilities for species which were observed only rarely.”  A common goal in network ecology is to describe interaction frequencies by using niche-based proxies (phylogeny, traits, etc). If it was a simply a matter that most networks had the same degree distribution, then predicting networks would be much easier than it is. In fact, we are really bad at predicting pairwise interactions (see Olito and Fox), even when we get the emergent structure of the network correct. Yet empirical bayes assumes a kind of commonality among all networks to help narrow the posteriors. I think the authors need to speak to this concern.

		\end{refquote}

		\textbf{R:} We do not share the Reviewer's negative view of empirical Bayes. Perhaps this is a matter of perspective: in a system where a great deal of information about the traits structuring interactions is known and trait values are available for most species, it is absolutely best to use a trait-based model to predict interactions. This is why we began our section on informative priors with trait-based models and only introduce empirical and approximate Bayes as solutions where trait information/models are lacking or where over-fitting is a strong concern. Most of the authors of this manuscript are more used to such relatively information-poor systems. In these cases, we agree that empirical Bayes is not perfect but argue that it is much better than nothing (or an uninformative prior, as long as a somewhat similar network is available).

@@

		Moreover, while we agree that empirical Bayes does not take differences in niche proxies into account, this may not be crucial in all studies. Researchers are often more interested in the overall structure of the network or in some emergent property such as stability than in the relationships between niche proxies and interaction frequencies. In such cases, the number of additional interactions we expect to be observed with further sampling is likely more important than the exact identity of those interactions, and an empirical Bayes approach is reasonable if a good trait-based model and trait information or not available.

		Finally, we disagree with the Reviewer's contention that ``empirical Bayes assumes a kind of commonality among all networks". This is not so; empirical Bayes only assumes some commonality between the focal network and those used to construct the prior. We have revised the text to emphasize the need to use similar networks to construct an empirical Bayes prior and note that \emph{Appendix S7} demonstrates the pitfalls of basing an empirical Bayes prior on a network with little commonality with the focal system.

@@

		\begin{quotation}

		Note that here the prior distribution will be the same for all interaction probabilities. It is also possible to interpret other network properties (e.g., species degrees) in a similar network(s) as interaction probabilities and use these to create a prior (worked example in \emph{Appendix S4}).

		\end{quotation}

		Finally, we discuss caveats to the empirical Bayes approach in lines 272-281, including the possibility that priors based on substantially different systems will be misleading (with reference to \emph{Appendix S7}) and that published data may reflect under-sampled data:

### 2019-03-04 manuscript/reply_to_reviewers_2.tex

		\begin{refquote}

			Figure 3 is fun and deserves a bit more central place in the discussion.

		\end{refquote}

		\textbf{R:} We thank the Reviewer for his complement to Figure 3. Although we wish to keep any additions as brief as possible, we are happy to add a short reference to Figure 3 in the discussion as part of an example of how to use our simple Bayesian framework when thinking about insects moving into new communities (e.g., as climate changes). We hope that this puts the figure in a more central place, as intended. We also now include a similar figure demonstrating the effect of a change in prior on the posterior pdf's obtained for the \emph{Salix}-galler networks and hope that this also helps to emphasise these results.

@@

		\end{quotation}

	2. Emphasise the drawbacks of empirical Bayes

		\begin{refquote}

			I think the authors miss one key discomfort with empirical Bayes. It seems vaguely anti-ecology, because it assumes that we can L289 : “use information from abundant and easily-sampled species to predict interaction probabilities for species which were observed only rarely.”  A common goal in network ecology is to describe interaction frequencies by using niche-based proxies (phylogeny, traits, etc). If it was a simply a matter that most networks had the same degree distribution, then predicting networks would be much easier than it is. In fact, we are really bad at predicting pairwise interactions (see Olito and Fox), even when we get the emergent structure of the network correct. Yet empirical Bayes assumes a kind of commonality among all networks to help narrow the posteriors. I think the authors need to speak to this concern.

@@

		Moreover, while we agree that empirical Bayes does not take differences in niche proxies into account, this may not be crucial in all studies. Researchers are often more interested in the overall structure of the network or in some emergent property such as stability than in the relationships between niche proxies and interaction frequencies. In such cases, the number of additional interactions we expect to be observed with further sampling is likely more important than the exact identity of those interactions, and an empirical Bayes approach is reasonable if a good trait-based model and trait information or not available.

		Finally, we disagree with the Reviewer's contention that ``empirical Bayes assumes a kind of commonality among all networks". This is not so; empirical Bayes only assumes some commonality between the focal network and those used to construct the prior. We have revised the text to emphasise the need to use similar networks to construct an empirical Bayes prior and note that \emph{Appendix S7} demonstrates the pitfalls of basing an empirical Bayes prior on a network with little commonality with the focal system.

@@

		\end{quotation}

	3. Include material from \emph{Appendix 7} in the main text

		\begin{refquote}

			My final lingering concern is that something like appendix 7 seems important and should be included in the text? Was this done because of space limitations? I think the authors can appeal to the editor that the paper merits the extra space. As a compromise there could to be at least a paragraph on what can go wrong and what that looks like. I think a figure would be useful here. If the informed/empirical priors are too narrow, the data will not really move the posterior. The authors certainly acknowledge it in various places, but it needs to be clearer to the reader.

@@

		\end{refquote}

		\textbf{R:} This material was indeed placed in an appendix because of space constraints. We have added a figure showing the posterior distributions obtained from both priors and a paragraph summarising \emph{Appendix 7}. As we do not wish to be too demanding with regard to extra space (and our manuscript was essentially at the standard length limit after the last revision), we hope that these smaller additions will be enough to emphasise the potential for spurious results after poor prior choices. Also note that the results for the galler-parasitoid community were somewhat similar. In order to more clearly demonstrate the potentially large effects of prior choice, we therefore showcase the results from \emph{Salix}-galler networks in this paragraph.

### 2019-03-04 manuscript/reply_to_reviewers_2.tex

	\begin{refquote}

		Thank you for the extensive and detailed reply to the comments of the referees. Your revised manuscript was sent back to one of the referees who reviewed the original version. As you will see, he is very positive about the revised version and has now signed his review. The only outstanding issue is the addition of some discussion about how to detect potential problems with the priors. I agree with this suggestion as one of the most common criticisms of Bayesian methods is precisely the use of priors.

	\end{refquote}

	\textbf{R:} We appreciate the additional comments from the Referee, and agree that the definition of adequate priors can be quite a challenge. We are therefore quite happy to add to our discussion about how to detect potential problems. Note, however, that after making the additions requested by the previous round of reviews our manuscript was very close to the standard word limit for \textbf{Methods in Ecology and Evolution}. To meaningfully expand our discussion of how to detect problems with priors we must therefore ask for more space-- a solution also explicitly recommended by the Reviewer. We have kept our addition to the discussion as succinct as possible, and the revised manuscript is now approximately 7433 words. We hope that this additional length is justified by our efforts to address the Reviewer's remaining concerns.

@@

		Sincerely,

		Ben Weinstein

	\end{refquote}

	We thank the Reviewer for his previous and further feedback and address each point (prefaced by \textbf{R:}) below.

@@

		\begin{refquote}

			I think the authors miss one key discomfort with empirical Bayes. It seems vaguely anti-ecology, because it assumes that we can L289 : “use information from abundant and easily-sampled species to predict interaction probabilities for species which were observed only rarely.”  A common goal in network ecology is to describe interaction frequencies by using niche-based proxies (phylogeny, traits, etc). If it was a simply a matter that most networks had the same degree distribution, then predicting networks would be much easier than it is. In fact, we are really bad at predicting pairwise interactions (see Olito and Fox), even when we get the emergent structure of the network correct. Yet empirical Bayes assumes a kind of commonality among all networks to help narrow the posteriors. I think the authors need to speak to this concern.

		\end{refquote}

		\textbf{R:} We appreciate but do not fully share the Reviewer's negative view of empirical Bayes. Perhaps this is a matter of perspective: in a system where a great deal of information about the traits structuring interactions is known and trait values are available for most species, trait-based models are indeed often a good way to predict interactions (although far from fool-proof, as the Reviewer points out). This is why we began our section on informative priors with trait-based models and only introduce empirical and approximate Bayes as solutions where trait information/models are lacking or where over-fitting is a strong concern. Most of the authors of this manuscript are more used to such relatively information-poor systems. In these cases, we agree that empirical Bayes is not perfect, but argue that it is much better than nothing (or an uninformative prior, as long as a somewhat similar network is available).

		[[Currently one vote for removing this paragraph and one vote for keeping it and strengthening the point]]
		The Jeffreys prior gives a maximum likelihood estimate of 0.5 for the probability of an interaction between two species which have never been observed together. Based on our understanding of ecological systems, this seems highly unlikely. Certainly species differ in many ways, but they also do have some similarities, such as limited foraging time and somewhat consistent morphologies. Because of behavioural, morphological, or temporal limits to the number of partners with which a given species can interact, most ecological networks are sparse and most well-sampled species interact with only a small subset of their potential partners (extreme generalists in highly nested mutualistic networks being a notable exception). We can therefore reasonably expect that a species entering a new habitat, for example, will interact with far fewer than half of the available partners. If we do not know which traits strongly affect interactions involving the focal species, or if those traits are not known for all of the species in the network, we cannot use a trait-based model to predict interaction frequencies. This leaves empirical and approximate Bayes as the only viable options, unless we are willing to revert to the assumption that unobserved interactions never occur. Certainly empirical Bayes is not perfect, but we are confident that it is still gives more reasonable expectations than ignoring the ecological networks literature.

		Moreover, while we agree that empirical Bayes does not take differences in niche proxies into account, this may not be crucial in all studies. Researchers are often more interested in the overall structure of the network or in some emergent property such as stability than in the relationships between niche proxies and interaction frequencies. In such cases, the number of additional interactions that we expect to be observed with further sampling is likely more important than the exact identity of those interactions, and an empirical Bayes approach is reasonable if a good trait-based model and trait information or not available.

		Finally, we disagree with the Reviewer's contention that ``empirical Bayes assumes a kind of commonality among all networks". This is not so; empirical Bayes only assumes some commonality between the focal network and those used to construct the prior. We have revised the text to emphasise the need to use similar networks to construct an empirical Bayes prior and note that \emph{Appendix S7} demonstrates the pitfalls of basing an empirical Bayes prior on a network with little commonality with the focal system. [[The importance of using prior information from similar networks is a small but very important point that refers back to the core of our argument: if informatiom about a focal system is limited, then closing our eyes to what is known about similar networks and assuming \emph{a priori} that an unobserved interaction in any system is either equally (and rather highly) likely or equally impossible seems more than ``vaguely anti-ecology". After all, ecology advances by building upon previous insights and not by ignoring them in an effort to be ``conservative". It is essential to realise that ``uninformative" priors can lead to very biased results, just as poorly-chosen informative priors can. ]][[This last portion strengthens the point in paragraph 2 but may be too much.]]

@@

		\textbf{R:} This material was indeed placed in an appendix because of space constraints. We have added a figure showing the posterior distributions obtained from both priors and a paragraph summarising \emph{Appendix 7}. As we do not wish to be too demanding with regard to extra space (and our manuscript was essentially at the standard length limit after the last revision), we hope that these smaller additions will be enough to emphasise the potential for spurious results after poor prior choices. Also note that the results for the galler-parasitoid community were somewhat similar. In order to more clearly demonstrate the potentially large effects of prior choice, we therefore showcase the results from \emph{Salix}-galler networks in this paragraph.

		Lines 390-405:

		\begin{quotation}

		  Note that, as in all Bayesian analyses, our results do depend on the prior chosen. To demonstrate this, we repeated our analyses using a prior derived from a study of gallers found on several genotypes of \emph{Salix hookeriana} and the parasites which emerged from them (\citealp{Barbour2016,Barbour2016Dryad}; \emph{Appendix S7}). Although the study system is similar to that in~\citet{Kopelke2017}, the network is quite different due to using different genotypes of a single \emph{Salix} species rather than several \emph{Salix} species as the basis for sampling. While this had a relatively small effect on our expectations for the galler-parasitoid community (\emph{Appendix S7}), the prior based on \citep{Barbour2016} resulted in very high probabilities of interaction between \emph{Salix} and galler pairs that were not observed interacting (Fig. 4). No amount of additional sampling would allow us to conclude that a given \emph{Salix}-galler pair did not interact with a threshold interaction probability of 0.01. This is reasonable for the situation described in~\citet{Barbour2016}, as it is very likely that gallers which can interact with one \emph{S. hookeriana} genotype can interact with most others, but is not reasonable for the more diverse community in~\citet{Kopelke2017}. This demonstrates the importance of conducting a ``sanity check" on the posterior distribution obtained from any given prior.

### 2019-03-04 manuscript/CI_interactions.tex
removed comments
added Tomas comments to reply, caught a mis-formatted quote in main text

  Note that, as in all Bayesian analyses, our results do depend on the prior chosen. To demonstrate this, we repeated our analyses using a prior derived from a study of gallers found on several genotypes of \emph{Salix hookeriana} and the parasites which emerged from them (\citealp{Barbour2016,Barbour2016Dryad}; \emph{Appendix S7}). Although the study system is similar to that in~\citet{Kopelke2017}, the network is quite different due to using different genotypes of a single \emph{Salix} species rather than several \emph{Salix} species as the basis for sampling. While this had a relatively small effect on our expectations for the galler-parasitoid community (\emph{Appendix S7}), the prior based on \citep{Barbour2016} resulted in very high probabilities of interaction between \emph{Salix} and galler pairs that were not observed interacting (Fig.~\ref{prior_comparison}). No amount of additional sampling would allow us to conclude that a given \emph{Salix}-galler pair did not interact with a threshold interaction probability of 0.01. This is reasonable for the situation described in~\citet{Barbour2016}, as it is very likely that gallers which can interact with one \emph{S. hookeriana} genotype can interact with most others, but is not reasonable for the more diverse community in~\citet{Kopelke2017}. This demonstrates the importance of conducting a "sanity check" on the posterior distribution obtained from any given prior.

@@

  % \begin{figure}[h!]
    %   \caption{Imperfect detection of interactions increases the number of samples required to be confident that an interaction never occurs. Assume that two species cannot interact (i.e., $\lambda=0$) and that the number of observed interactions follows a binomial distribution depending upon the number of observed interactions $k$ and observed co-occurrences $n$. The maximum likelihood estimate (MLE) for $\lambda_{ij}$ is $\lambda_{MLE}=\frac{k}{n}$. Note that $\lambda_{ij}$ is not a point estimate but rather a random variable with an unknown distribution. This means that if $k=0$ in a given sample, this does not necessarily imply that the two species will never interact. Rather, $k=0$ implies that `no interaction' is the most likely outcome when the species do co-occur but there is nonetheless some chance that the two species \emph{could} interact.\\
    %   Importantly, we can estimate the variance of $\lambda_{ij}$ as well as the MLE; available methods include the \emph{Wilson score interval}, the \emph{Clopper-Pearson interval}, and the \emph{Agresti-Coull interval} (for details, see [\citealp{Brown2001}]). The distribution of the MLE for $\lambda_{ij}$ thus obtained is a key part of the Bayesian distribution for $\lambda_{ij}$.
    %   Here we show the upper bound (solid black line) of a 95\% Clopper-Pearson true credible interval for the interaction probability $\lambda_{ij}$ for a pair of species that has been observed co-occurring $n$ times but never interacting ($k=0$). The upper limit of the credible interval only reaches 0.1 (dashed, red line) with 35 observations. Thus, adding more observations is useful in controlling uncertainty, but the number of observations added must be very high unless we have some other reason to believe that an interaction cannot occur (i.e., a strong prior).}
    %   \label{upper_limits}
    %   \begin{center}
      % \includegraphics*[width=.8\textwidth]{figures/upper_limit_DG.eps}
    %   % \includegraphics*[width=.8\textwidth]{Figure_3.eps}
    %   \end{center}
    %   \end{figure}

  \begin{figure}[h!]
    \caption{Posterior distributions for the interaction probability ($\lambda_{ij}$) for two species that have not been observed interacting ($k_{ij}=0$), based on a single site (Zillis in Graub\"{u}nden, Switzerland) from~\citet{Kopelke2017}. Posterior distributions (curves) and 95\% credible intervals (lines at top of panel) for $\lambda_{ij}$ narrow and approach zero as the number of observed co-occurrences ($n_{ij}$) increases. Diamonds indicate the maximum likelihood estimator for the mean probability of interaction.
    Dashed lines indicate threshold probabilities of 0.01, 0.05, and 0.1. The number of samples required to obtain a 95\% credible interval below each threshold increases rapidly. It takes just over 20 observed co-occurrences to be 95\% confident that $\lambda_{ij}<0.10$ but over 100 co-occurrences to be 95\% confident that $\lambda_{ij}<0.01$.}
    \label{Salix_pdfs_cdfs}
    \begin{center}
    \includegraphics*[width=.8\textwidth]{figures/GP_pdfs_increasing_N_Zillis.eps}
    % \includegraphics*[width=.8\textwidth]{Figure_4.eps}
    \end{center}
    \end{figure}

  \begin{figure}[h!]
    \caption{Posterior distributions for the interaction probability ($\lambda_{ij}$) for two species that have not been observed interacting ($k_{ij}=0$), A) based on a single site (Zillis in Graub\"{u}nden, Switzerland) from~\citet{Kopelke2017} and B) based on a network describing interactions between gallers and several genotypes of \emph{Salix hookeriana} from~\citet{Barbour2016}. Posterior distributions (curves) and 95\% credible intervals (lines at top of panel) for $\lambda_{ij}$ narrow and approach zero as the number of observed co-occurrences ($n_{ij}$) increases. Diamonds indicate the maximum likelihood estimator for the mean probability of interaction.
    Dashed lines indicate threshold probabilities of 0.01, 0.05, and 0.1. Note that the prior obtained from~\citet{Barbour2016} constrains the MLE  of $\lambda_{ij}$ above 0.01.}
    \label{prior_comparison}
    \begin{center}
    \includegraphics*[width=.8\textwidth]{figures/SG_pdfs_prior_comparison.eps}
    % \includegraphics*[width=.8\textwidth]{Figure_4.eps}
    \end{center}
    \end{figure}

  % \begin{figure}[h!]
  %   \caption{Mean connectance, links per galler, nestedness (NODF), and links per natural enemy for networks assembled using posterior distributions based on a single site (Zillis in Graub\"{u}nden, Switzerland) from~\citet{Kopelke2017} occupied a narrow range. To obtain distributions of network properties, we created 100 ``posterior-sampling" networks and then, for each of these, created 100 ``filtered" networks by randomly sampling 50\%-99\% of the interactions included in the posterior-sampling network. This simulates imperfect detection of interactions in the field. Each point represents the mean (+/- SD) network property (e.g., connectance) obtained from a set of 100 filtered networks. The filtered networks cover a broader range of network properties than the posterior-sampling networks but the value of each property decreases as the strength of the detection filter increases. Values in the original web are indicated by dashed lines.}
  %   \label{posterior_webs}    
  %   \begin{center}
  %   % \includegraphics[width=.8\textwidth]{figures/GP_posterior_properties_Zillis.eps}
  %   % \includegraphics[width=.8\textwidth]{Figure_5.eps}
  %   \end{center}
  %   \end{figure}

>>>>

  Note that, as in all Bayesian analyses, our results do depend on the prior chosen. To demonstrate this, we repeated our analyses using a prior derived from a study of gallers found on several genotypes of \emph{Salix hookeriana} and the parasites which emerged from them (\citealp{Barbour2016,Barbour2016Dryad}; \emph{Appendix S7}). Although the study system is similar to that in~\citet{Kopelke2017}, the network is quite different due to using different genotypes of a single \emph{Salix} species rather than several \emph{Salix} species as the basis for sampling. While this had a relatively small effect on our expectations for the galler-parasitoid community (\emph{Appendix S7}), the prior based on \citep{Barbour2016} resulted in very high probabilities of interaction between \emph{Salix} and galler pairs that were not observed interacting (Fig.~\ref{prior_comparison}). No amount of additional sampling would allow us to conclude that a given \emph{Salix}-galler pair did not interact with a threshold interaction probability of 0.01. This is reasonable for the situation described in~\citet{Barbour2016}, as it is very likely that gallers which can interact with one \emph{S. hookeriana} genotype can interact with most others, but is not reasonable for the more diverse community in~\citet{Kopelke2017}. This demonstrates the importance of conducting a ``sanity check" on the posterior distribution obtained from any given prior.

@@

  \begin{figure}[h!]
    \caption{Posterior distributions for the interaction probability ($\lambda_{ij}$) for two species that have not been observed interacting ($k_{ij}=0$), based on a single site (Zillis in Graub\"{u}nden, Switzerland) from~\citet{Kopelke2017}. Posterior distributions (curves) and 95\% credible intervals (lines at top of panel) for $\lambda_{ij}$ narrow and approach zero as the number of observed co-occurrences ($n_{ij}$) increases. Diamonds indicate the maximum likelihood estimator for the mean probability of interaction.
    Dashed lines indicate threshold probabilities of 0.01, 0.05, and 0.1. The number of samples required to obtain a 95\% credible interval below each threshold increases rapidly. It takes just over 20 observed co-occurrences to be 95\% confident that $\lambda_{ij}<0.10$ but over 100 co-occurrences to be 95\% confident that $\lambda_{ij}<0.01$.}
    \label{Salix_pdfs_cdfs}
    \begin{center}
    \includegraphics*[width=.8\textwidth]{figures/GP_pdfs_increasing_N_Zillis.eps}
    % \includegraphics*[width=.8\textwidth]{Figure_3.eps}
    \end{center}
    \end{figure}

  \begin{figure}[h!]
    \caption{Posterior distributions for the interaction probability ($\lambda_{ij}$) for two species that have not been observed interacting ($k_{ij}=0$), A) based on a single site (Zillis in Graub\"{u}nden, Switzerland) from~\citet{Kopelke2017} and B) based on a network describing interactions between gallers and several genotypes of \emph{Salix hookeriana} from~\citet{Barbour2016}. Posterior distributions (curves) and 95\% credible intervals (lines at top of panel) for $\lambda_{ij}$ narrow and approach zero as the number of observed co-occurrences ($n_{ij}$) increases. Diamonds indicate the maximum likelihood estimator for the mean probability of interaction.
    Dashed lines indicate threshold probabilities of 0.01, 0.05, and 0.1. Note that both priors constrain the MLE of $\lambda_{ij}$ above 0.1 and that the posterior probability distributions become similar for very high $n$.}
    \label{prior_comparison}
    \begin{center}
    \includegraphics*[width=.8\textwidth]{figures/SG_pdfs_prior_comparison.eps}
    % \includegraphics*[width=.8\textwidth]{Figure_4.eps}
    \end{center}
    \end{figure}
